\bf An Examination of Sport Climbing Competition Format and Scoring System

### 2021-09-05 manuscript/Rmd/draft.Rmd
New abstract

abstract: |
  The purpose of this paper is to investigate the controversial competition format and scoring system of sport climbing, which is one of the sports making its debut on the Olympics stage at Tokyo 2020. Climbing Olympians are required to participate in three disciplines: speed climbing, bouldering, and lead climbing for only a single set of medals, and a multiplied score of climbers' rankings in all three disciplines will be used to determine the overall standings. This work finds great evidence from historical data and simulations that the current competition format is putting speed climbers at a disadvantage. Furthermore, it is shown that the product of ranks scoring system violates the independence of irrelevant alternatives property.

@@

>>>>

abstract: |
  Sport climbing, which made its Olympic debut at the 2020 Summer Games, generally consists of three separate disciplines: speed climbing, bouldering, and lead climbing.  However, the International Olympic Committee (IOC) only allowed one set of medals per gender for sport climbing at Tokyo 2020. As a result, the governing body of sport climbing, rather than choosing only one of the three disciplines to include in the Olympics, decided to create a competition combining all three disciplines. In order to determine a winner, a combined scoring system was created using the product of the ranks across the three disciplines to determine an overall score for each climber. In this work, the sport climbing scoring system is evaluated through simulations to investigate its general features such as the advancement probabilities given certain placements and the sensitivity of the scoring system to small changes. In addition, analyses of historical climbing contest results are presented and real examples of violations of the independence of irrelevant alternatives are illustrated. Finally, this work finds evidence that the current competition format is putting speed climbers at a disadvantage.

@@

gold medal

order of medalists

3 medalists

any rankings

Structure:

sensitivity to:

results

tournament format

### 2021-09-05 manuscript/Rmd/draft.tex
New abstract

\bigskip
\begin{abstract}
The purpose of this paper is to investigate the controversial
competition format and scoring system of sport climbing, which is one of
the sports making its debut on the Olympics stage at Tokyo 2020.
Climbing Olympians are required to participate in three disciplines:
speed climbing, bouldering, and lead climbing for only a single set of
medals, and a multiplied score of climbers' rankings in all three
disciplines will be used to determine the overall standings. This work
finds great evidence from historical data and simulations that the
current competition format is putting speed climbers at a disadvantage.
Furthermore, it is shown that the product of ranks scoring system
violates the independence of irrelevant alternatives property.
\end{abstract}

>>>>

\bigskip
\begin{abstract}
Sport climbing, which made its Olympic debut at the 2020 Summer Games,
generally consists of three separate disciplines: speed climbing,
bouldering, and lead climbing. However, the International Olympic
Committee (IOC) only allowed one set of medals per gender for sport
climbing at Tokyo 2020. As a result, the governing body of sport
climbing, rather than choosing only one of the three disciplines to
include in the Olympics, decided to create a competition combining all
three disciplines. In order to determine a winner, a combined scoring
system was created using the product of the ranks across the three
disciplines to determine an overall score for each climber. In this
work, the sport climbing scoring system is evaluated through simulations
to investigate its general features such as the advancement
probabilities given certain placements and the sensitivity of the
scoring system to small changes. In addition, analyses of historical
climbing contest results are presented and real examples of violations
of the independence of irrelevant alternatives are illustrated. Finally,
this work finds evidence that the current competition format is putting
speed climbers at a disadvantage.
\end{abstract}
